---
title: "Linear Models"
author: "ak5357"
date: "2024-11-07"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
```

## Data Cleaning

```{r}
data("nyc_airbnb")

nyc_airbnb = nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood
  ) |> 
  filter(
    borough != "Staten Island"
  ) |> 
  select(price, stars, borough, neighborhood, room_type) |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

## Fit some models

Start relatively simple

```{r}
# Linear model
fit = lm(price ~ stars, data = nyc_airbnb)

# Get short summary
summary(fit)

# What all does it tell you?
names(fit)
names(summary(fit))

# Get coefficients
coef(fit)

# Make values easier to access, as a tibble
broom::tidy(fit)

# Using piping
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value)
```


Try a bit more complex.

* For categorical variable, first value (Bronx) becomes reference and the rest get compared to that.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  knitr::kable(digits = 3)

```


## Some diagnostics

Most diagnostics use residuals.

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = resid)) +
  geom_histogram()

```

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-200, 500)

```


Residuals against stars.

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```


Residuals against fitted values.

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```


## Hypothesis testing

For single coefficients, look at the table.
```{r}
fit |> broom::tidy()
```

When adding more than one coefficient, use "ANOVA".

* This only works if you're using _nested_ models. Otherwise, ANOVA doesn't help you.
```{R}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

Look at both.
```{r}
fit_null |> 
  broom::tidy()

fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Do effects differ across boroughs?

First, use a lot of interactions.

```{r}
nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) |> 
  broom::tidy()
```

Could just fit separate models.

```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

Get fancy and use list columns.

```{r}
nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```


What about room types across manhattan neighborhoods?

```{r}
nyc_airbnb |> 
  filter(
    borough == "Manhattan",
    neighborhood == "Chinatown") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
  
  
nyc_airbnb |> 
  filter(
    borough == "Manhattan",
    neighborhood == "Chelsea") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

```{r}
manhattan_neighborhood_fit_df = 
  nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  nest(data = -(borough:neighborhood)) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) |> 
  select(neighborhood, results) |> 
  unnest(results)
```


Look at the effect of room type

```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = term, y = estimate)) +
  geom_boxplot()
```


```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  filter(estimate > 0)

nyc_airbnb |> 
  filter(
    neighborhood == "NoHo",
    room_type == "Shared room"
    )


```





